Unsupervised training of maximum-entropy models for lexical selection in rule-based machine translation

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
some trimming
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58783 72bbbca6-d526-0410-a7d9-f06f51895060

\begin{abstract}
  This article presents a method of training maximum-entropy models to
  perform lexical selection in a rule-based machine translation
  system. The training method described is unsupervised; that is, it
  does not require any annotated corpus. The method uses
  source-language monolingual corpora, an existing machine translation
  (MT) system and a statistical target-language model. Using the
  existing MT system, the sentences in the source-language corpus are
  translated in all possible ways according to the different
  translation equivalents in the bilingual dictionary of the system.
%  taking into account all of the lexical selection possibilities.
  These translations are then scored on the target-language model and
  the scores are normalised to provide fractional counts for training
  a source-language maximum-entropy lexical-selection model. We show
  that this model can perform equally well, or better, than using the
  target-language model directly for lexical selection, at a
  substantially reduced computational cost.
\end{abstract}

@@

\subsection{Prior work}
\comment{This section should be ironed out a bit.}  In their
pioneering work, \cite{dagan1994word} used the term
\emph{word sense disambiguation} to refer to what is actually target
word selection in machine translation; they used a parser to identify
some syntactic relations such as subject--object or subject--verb.
After generating all the possible translations for a given
input sentence using an ambiguous bilingual dictionary, they extract the syntactic
tuples from the target language and count the frequency in a previously-trained target-language 
model of tuples.  They use maximum-likelihood estimation to calculate the probability of a given 
target language tuple being the translation of a given source language tuple, with
an automatically determined threshold for confidence.

Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
on the specific problem of lexical selection in IBM-style word-based
statistical machine translation. Other authors
\citep{melero07a} have used target-language models to rank %@@@ carbonell06
the translations resulting from all possible combinations of lexical
selections.  Nowadays, in state-of-the-art phrase-based statistical
machine translation \citep{koehn2009statistical}, lexical selection is
taken care of by a combination of the translation model and the
language model. The translation model provides probabilities of
translation between words or word sequences (often referred to as
\emph{phrases}) in the source and target language. The target-language
model provides probabilities of word sequences in the target language.
As regards rule-based machine translation, \cite{sanchez07a} proposed 
to use information from the target language to train a statistical 
lexical-selection module but did not provide any experiments. 
\todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
using a bilingual corpus.  More recently, \cite{tyers12a} presented a
method of lexical selection for rule-based machine translation based
on rules which select or remove translations in fixed-length contexts,
along with a training method for learning the rules from a
word-aligned parallel corpus. The principal disadvantage with this
method is that it relies on a parallel corpus for the language pair,
something which is not always available.

>>>>

\begin{abstract}
  This article presents a method of training maximum-entropy models to
  perform lexical selection in a rule-based machine translation
  system. The training method described is unsupervised; that is, it
  does not require any annotated corpus. The method uses
  source-language monolingual corpora, an existing machine translation
  (MT) system and a statistical target-language model. Using the
  existing MT system, the sentences in the source-language corpus are
  translated in all possible ways according to the different
  translation equivalents in the bilingual dictionary of the system.
  These translations are then scored on the target-language model and
  the scores are normalised to provide fractional counts for training
  a source-language maximum-entropy lexical-selection model. We show
  that this model can perform equally well, or better, than using the
  target-language model directly for lexical selection, at a
  substantially reduced computational cost.
\end{abstract}

@@

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
In their
pioneering work, \cite{dagan1994word} used the term
\emph{word sense disambiguation} to refer to what is actually target
word selection in machine translation; they used a parser to identify
some syntactic relations such as subject--object or subject--verb.
After generating all the possible translations for a given
input sentence using an ambiguous bilingual dictionary, they extract the syntactic
tuples from the target language and count the frequency in a previously-trained target-language 
model of tuples.  They use maximum-likelihood estimation to calculate the probability of a given 
target language tuple being the translation of a given source language tuple, with
an automatically determined threshold for confidence.

Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
on the specific problem of lexical selection in IBM-style word-based
statistical machine translation. Other authors
\citep{melero07a} have used target-language models to rank %@@@ carbonell06
the translations resulting from all possible combinations of lexical
selections.  Nowadays, in state-of-the-art phrase-based statistical
machine translation \citep{koehn2009statistical}, lexical selection is
taken care of by a combination of the translation model and the
language model. The translation model provides probabilities of
translation between words or word sequences (often referred to as
\emph{phrases}) in the source and target language. The target-language
model provides probabilities of word sequences in the target language.
As regards rule-based machine translation, \cite{sanchez07a} proposed 
to use information from the target language to train a statistical 
lexical-selection module but did not provide any experiments. 
%@@@ \todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
using a bilingual corpus.  More recently, \cite{tyers12a} presented a
method of lexical selection for rule-based machine translation based
on rules which select or remove translations in fixed-length contexts,
along with a training method for learning the rules from a
word-aligned parallel corpus.

% The principal disadvantage with this
% method is that it relies on a parallel corpus for the language pair,
% something which is not always available.
